---
title: "Lab 5"
author: "Kipras Varkala"
date: "2/17/2022"
output: html_document
---
# Lab 5

## Lab 5A
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
lotr  <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/lotr_tidy.csv") %>% 
  rename(Species = Race)

head(lotr)
```

## 1. Would you say this data is in tidy format?

Yes, I would say this data is in tidy format.


### 2. Widen the data so that we see the words spoken by each species, by putting species as its own column.

```{r Widen}
(lotr_wide <- lotr %>% 
  pivot_wider(id_cols = -c(Species, Words), 
              names_from = Species, 
              values_from = Words))
```
### 3. Re-lengthen the wide LOTR data from Question 2 above.

```{r Lengthen}
lotr_wide %>%
  pivot_longer(cols = c(Elf : Man), 
               names_to  = "Species", 
               values_to = "Words")
```

## Lab 5B

```{r setup 2, include=FALSE}
set.seed(123)
missing_w2_parent <- sample(1:500, 30)
missing_w2_child <- c(missing_w2_parent[1:5], sample(1:500, 25))
family  <- read_csv(
  "https://raw.githubusercontent.com/bwiernik/progdata/main/inst/tutorials/data/family_data.csv"
) |> 
  mutate(
    across(
      starts_with("w2") & contains("parent"),
      ~ ifelse(family_id %in% missing_w2_parent, NA_real_, .x)
    ),
    across(
      starts_with("w2") & contains("child"),
      ~ ifelse(family_id %in% missing_w2_child, NA_real_, .x)
    )
  )

family |> 
  DT::datatable(rownames = FALSE)
```

### 1. Reshape the dataset to a “longer” format. Make each row 1 score. 

```{r Reshape}
family_longest <- family %>%
  pivot_longer(
  cols = w1_comm_parent:w2_satis_child,
  names_to = c("wave", "measure", "family_member"), 
  names_sep = "_",
  values_to = "score"
) %>% 
  mutate(wave = stringr::str_replace(wave, "w", "")) %>%
  mutate(wave = as.integer(wave))

print(family_longest)
```

### 2. Reshape the dataset to a “longer” format. Make each row 1 person. 

```{r Reshape 2}
family_long <- family %>%
  pivot_longer(
    cols = -family_id,
    names_to = c("wave", ".value", "family_member"), 
    names_sep = "_"
  )

print(family_long)
```

### 3. Some families are missing wave 2 data for parent, child, or both. Which families are missing wave 2 data for at least one person?

```{r}
family_longest %>%
  filter(wave == 2 & is.na(score))
```

### Is it easier to easier to find the missing data in the wide or long format?

Long formats allow for easier identification of missing values as every row is represented by one person. 