---
title: "Lab 1 Penguin Acitivity"
author: "Kipras Varkala"
date: "1/20/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(palmerpenguins)
```

```{r load data}
pengy <- palmerpenguins::penguins
```


# Exploring Functions:

>Summary function

```{r summary function}
summary(pengy)

#Provides summary statistics about the entire data set
```

Using the summary function, we can see the counts of different groups and the basic descriptive statistics of the numeric columns (mean, median, mode, etc.).

>Head function

```{r head function}
head(pengy)

#Displays the first 6 rows of the data set. You can use the tail() function to display the last 6 rows as well. 
```

Using the head function, we can see a preview of the data set by looking at the first 6 rows. 

>Structure function

```{r structure function}
str(pengy)

#Outputs the structure of each column
```

Using the structure function, we can see that not all columns have the sae structure. The data set has columns in factor, numeric, and integer formats.

>Dimension function

```{r dimension function}
dim(pengy)

#Shows the dimensions of the data set: # of rows/# of columns
```

Using the dimension function we can see that there are 344 rows and 8 columns in the penguin dataset.

