---
title: "lab-02-plastic-waste"
author: "Kipras Varkala"
date: "1/31/2022"
output: html_document
---
## Load packages and data

```{r setup, message=FALSE}
library(tidyverse)
```

```{r load-data, message=FALSE}
plastic_waste <- read.csv(here::here("data/plastic-waste.csv"))

head(plastic_waste)

dplyr::glimpse(plastic_waste)
```

## Exercises

### Exercise 1



```{r plastic-waste-continent}
ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap) +
  facet_wrap(~continent) +
  geom_histogram()
```

### Exercise 2

```{r plastic-waste-density}

ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap) +
  facet_wrap(~continent) +
  geom_density(aes(y = after_stat(count)))
```

```{r filled by countinent and semi-transparent}
ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap, 
      fill = continent) +
  facet_wrap(~continent) +
  geom_density(aes(y = after_stat(count)),
               alpha = .3) 
```


### Exercise 3

We define color and fill in aes because

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-violin}
ggplot(data = plastic_waste) +
  aes(x = plastic_waste_per_cap) +
  geom_histogram(binwidth = 0.2)
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r plastic-waste-mismanaged}
# insert code here
```

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

```{r plastic-waste-mismanaged-continent}
ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap) +
  facet_wrap(~continent) +
  geom_density(aes(y = after_stat(count)))
```

### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
# insert code here
```

## Pro-Tips

### Excercise 3

Try this :D

```
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()+
  geom_boxplot(width=.3, fill="green") +
  stat_summary(fun.y=median, geom="point") 
```

### Exercise 5 

Helpful reference:http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements


