---
title: "Lab 4"
author: "Kipras Varkala"
date: "04/13/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(datawizard)
library(tidyr)
library(readr)
library(gapminder)
library(ggplot2)
library(ggdist)
```

```{r load data}
dictionary <- readr::read_csv(here::here("data", "green-dictionary.csv"))
green_data <- readr::read_csv(here::here("data", "green-data.csv"))
```


## Questions


### 1. Inspect the item responses (e.g., with graphs or by summarizing distinct values). Is anything unusual?

```{r Inspect}
dictionary %>% 
  glimpse()

green_data %>% 
  glimpse()
```
-99 is not an intended rating option so we must change it to NA. 

```{r Double Check, Replace, Remove, and Reverse}
green_data %>% #check
  select(-id) %>% 
  unlist() %>% 
  unique()

green_data <- #replace
  green_data %>%
  mutate(
    across(
      -id,
      \(x) recode(x, "1" = 1, "2" = 2, "3" = 3, "4" = 4, "5" = 5, "-99" = NA_real_)
    )
  )

green_data %>% #double check
  select(-id) %>% 
  unlist() %>% 
  unique()

green_data <- #remove
  green_data %>%
  drop_na()

#reverse
reversed_key <- dictionary %>% 
  filter(Keying == -1) %>% 
  pull(Item)

green_data <- 
  green_data %>%
  mutate(
    across(
      all_of(reversed_key),
      \(x) recode(x, "1" = 5, "2" = 4, "3" = 3, "4" = 2, "5" = 1))
    )
```

### 2. Compute total scores for the four scales. Recode variables as needed.

```{r Total Scores}
raw_totals <-
green_data %>%
  rowwise() %>%
  mutate(
    id = id,
    green_total = mean(c_across(starts_with("green")), na.rm = TRUE),
    comp_total = mean(c_across(starts_with("comp")), na.rm = TRUE),
    intel_total = mean(c_across(starts_with("intel")), na.rm = TRUE),
    open_total = mean(c_across(starts_with("open")), na.rm = TRUE),
    student = factor(student),
    .before = everything()
  ) %>%
  ungroup() 

Totals_separate <- raw_totals %>%
   select(
     c(green_total : open_total),
     student,
   )

head(raw_totals)

summary(Totals_separate)
```

### 3. Rescale the variables so that they go from 0-100 instead of the original range. Name the recaled variables *_pomp.

```{r Rescale}
re_green_data <- 
  raw_totals %>% 
  mutate(
    across(
      green_total : open_total, 
      function(x) datawizard::change_scale(x, range(1,5), to = c(0, 100)), 
      .names = "{.col}_pomp"
    ), 
    .after = id
  )

head(re_green_data)
```

### 4. Make plots that illustrate the distributions of the 4 POMP-scored variables.

```{r Histograms}
attach(re_green_data)
par(mfrow=c(2,2))
hist(green_total_pomp)
hist(comp_total_pomp)
hist(intel_total_pomp)
hist(open_total_pomp)
```

### 5. Make scatterplots showing the relationships between green reputation and each personality trait. Include trend lines for students and non-students. What do these plots show?

```{r Scatterplots}
ggplot(re_green_data) +
  aes(x = green_total_pomp, y = comp_total_pomp, color = student)+
  geom_point() +
  geom_smooth() +
  geom_jitter() +
  labs(title = 'Green Reputation and Compassion')
ggplot(re_green_data) +
  aes(x = green_total_pomp, y = open_total_pomp, color = student)+
  geom_point() +
  geom_smooth() +
  geom_jitter() +
  labs(title = 'Green Reputation and Openness')
ggplot(re_green_data) +
  aes(x = green_total_pomp, y = intel_total_pomp, color = student)+
  geom_point() +
  geom_smooth() +
  geom_jitter() +
  labs(title = 'Green Reputation and Curiosity')
```
Green reputation and compassion: pretty similar slightly positve linear relationship with students scoring slightly higher. 
Green reputation and openness: aside from the outliers in non-student, both groups share a positive curvilinear relationship. 
Green reputation and compassion: curiosity appears to be positvely correlated with green reputation. 

### 6. Compare green reputation for students and non-students using a rainfall plot (bar + density + data points).

```{r Rainfall Plot}
ggplot(re_green_data) +
  aes(
    x = green_total_pomp, 
    y = student, 
    fill = student, 
    color = student
    ) + 
  geom_jitter(
    height = 0.3, 
    width = 0.3
    ) +
  geom_boxplot(
    alpha = 0.3, 
    width = 0.3, 
    color = 'black'
    ) +
  ggdist::stat_slab(
    position = position_nudge(y = 0.2),
    alpha = 0.3, 
    size = 0.5, 
    height = .6
    )
```

### 7. Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.

```{r Summary}
re_green_data %>%
  group_by(student) %>%
  summarize(
    across(
      c(green_total,comp_total,intel_total,open_total), 
      list(mean = mean, SD = sd, median = median, min = min, max = max)
      )
    )
```





