---
title: "Lab 4"
author: "Kipras Varkala"
date: "2/10/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
```

```{r load data}
dictionary <- readr::read_csv(here::here("C:/Users/kvpsy/GitHub/ProgDataClass/data", "green-dictionary.csv"))
green_data <- readr::read_csv(here::here("C:/Users/kvpsy/GitHub/ProgDataClass/data", "green-data.csv"))
```
```{r}
raw_totals <-
green_data %>%
  rowwise() %>%
  mutate(
    id = id,
    green_total = mean(c_across(starts_with("green")), na.rm = TRUE),
    comp_total = mean(c_across(starts_with("comp")), na.rm = TRUE),
    intel_total = mean(c_across(starts_with("intel")), na.rm = TRUE),
    open_total = mean(c_across(starts_with("open")), na.rm = TRUE),
    .before = everything()
  ) %>%
  ungroup() 

head(raw_totals)

#plotting

raw_totals %>%
  ggplot() %>%
  aes(x = )
```

Inspect the item responses (e.g., with graphs or by summarizing distinct values). Is anything unusual?

Compute total scores for the four scales. Recode variables as needed.

Rescale the variables so that they go from 0-100 instead of the original range. Name the recaled variables *_pomp.

Make plots that illustrate the distributions of the 4 POMP-scored variables.

Make scatterplots showing the relationships between green reputation and each personality trait. Include trend lines for students and non-students. What do these plots show?

Compare green reputation for students and non-students using a rainfall plot (bar + density + data points).

Compute a summary table of means, SDs, medians, minima, and maxima for the four total scores for students and non-students.





